### 2023-11-22 Presentation/main.tex
Init with Code from Lemire + Paper-Template from LIPIcs + Presentation-Template from s9770652

\documentclass[aspectratio=169]{gpresentation}

\RequirePackage{marvosym}

\graphicspath{{img/}}

\tikzset{
	legend/.style = {font=\boldmath, align=left},
	every path/.style = {line width=.57mm, line cap=round},
}

\subject{Seminar Algorithms for Big Data}
\title{Fast Random Integer Generation in an Interval}
\subtitle{Based on a paper of the same title by Daniel Lemire}
\author{Lukas Geis}
\supervisor{Supervised by Dr Manuel Penschuck}
\date{29th February 2024}
\institute{Algorithm Engineering (Prof. Dr Ulrich Meyer)}

\begin{document}
	\begin{frame}[plain, noframenumbering]
		\titlepage
	\end{frame}

	\begin{frame}{Table of Contents}
		\tableofcontents%[pausesections]
	\end{frame}
\end{document}