### 2024-02-08 Presentation/main.tex
java algorithm

\documentclass[aspectratio=169]{gpresentation}

\RequirePackage{marvosym}

\usepackage{mathtools}
\usepackage{nicefrac}
\newcommand{\Mod}[1]{\ \mathrm{mod}\ #1}

\newcommand{\GB}[1]{\textcolor{goetheblau}{#1}}
\newcommand{\ER}[1]{\textcolor{emorot}{#1}}
\newcommand{\DG}[1]{\textcolor{dunkelgruen}{#1}}
\newcommand{\BB}[1]{\textcolor{black}{#1}}

\newcommand{\RNW}{\ensuremath{\mathcal{R}_n^W}}

\usepackage{everysel}
\everymath{\color{goetheblau}}

\graphicspath{{img/}}

\tikzset{
	legend/.style = {font=\boldmath, align=left},
	every path/.style = {line width=.57mm, line cap=round},
}

\tikzstyle{overbrace text style}=[font=\small, above, pos=.5, yshift=3mm]
\tikzstyle{overbrace style}=[decorate,decoration={brace,raise=2mm,amplitude=3pt}]
\tikzstyle{underbrace style}=[decorate,decoration={brace,raise=2mm,amplitude=3pt,mirror},color=gray]
\tikzstyle{underbrace text style}=[font=\small, below, pos=.5, yshift=-3mm]

\usepackage{cancel}

\subject{Seminar Algorithms for Big Data}
\title{Fast Random Integer Generation in an Interval}
\subtitle{Based on a paper of the same title by Daniel Lemire}
\author{Lukas Geis}
\supervisor{Supervised by Dr. Manuel Penschuck}
\date{29th February 2024}
\institute{Algorithm Engineering (Prof. Dr. Ulrich Meyer)}

\begin{document}
	\begin{frame}[plain, noframenumbering]
		\titlepage
	\end{frame}

	\include{motivation}

	\begin{frame}{Table of Contents}
		\tableofcontents%[pausesections]
	\end{frame}

	\include{preliminaries}

	\include{unbiased}

	\include{lemire}

	\include{conclusion}
\end{document}