### 2024-02-08 Presentation/unbiased.tex

\section{Unbiased Algorithms}

\subsection{The OpenBSD Algorithm}
\begin{frame}{The OpenBSD Algorithm}
    \pause 
    \begin{itemize}[<+->]
        \item We shift the \ER{rejection interval} to the left:
        \onslide<+->{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\ER{\underbrace{0,1,\ldots,\RNW - 1}_{\text{$\RNW$ values}}},\underbrace{\overbrace{\RNW,\ldots,n - 1,0,\ldots,\RNW - 1}^{\text{$n$ values}},\ldots,\overbrace{\RNW,\ldots,n - 1,0,\ldots,\RNW - 1}^{\text{$n$ values}}}_{\text{$\left(2^W \div n\right) \cdot n$ values}}}^{\text{$2^W$ values}}
        \end{align*}}
        \item Generate a uniform random number $x \in [0,2^W)$ until $x \geq \RNW$
        \item Return $x \Mod n$
    \end{itemize}

    \begin{block}{Efficiency}<+->
        \onslide<+->{We require $2$ integer division operations:}
        \onslide<+->{one for computing $\RNW$}
        \onslide<+->{and one for computing $x \Mod n$.}
    \end{block}

\end{frame}

%%% Array Drawing with Tikz instead of Equation
\iffalse
\onslide<+->{\begin{center}
    \begin{tikzpicture}[scale = 0.7]
        \draw[draw = goetheblau] (0.0,0.0) rectangle ++(5.0,1.0);
        \draw[draw = goetheblau,overbrace style] (0.0,1.0) -- (5.0,1.0) node [overbrace text style] {\GB{$n$ values}};
        \draw[draw = goetheblau] (5.0,0.0) rectangle ++(5.0,1.0);
        \draw[draw = goetheblau,overbrace style] (5.0,1.0) -- (10.0,1.0) node [overbrace text style] {\GB{$n$ values}};
        \node at (10.5,0.5) {$\cdots$};
        \draw[draw = goetheblau] (11.0,0.0) rectangle ++(5.0,1.0);
        \draw[draw = goetheblau,overbrace style] (11.0,1.0) -- (16.0,1.0) node [overbrace text style] {\GB{$n$ values}};
        \draw[draw = emorot] (16.0,0.0) rectangle ++(3.0,1.0);
        \draw[draw = emorot,overbrace style] (16.0,1.0) -- (19.0,1.0) node [overbrace text style] {$\text{\ER{\RNW values}}$};
    \end{tikzpicture}
\end{center}}
\fi

\subsection{The Java Algorithm}
\begin{frame}{The Java Algorithm}
    \pause 
    \begin{itemize}
        \item<1-> Consider $x - \left(x \Mod n\right)$ for $x \in [0,2^W)$: \only<3>{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}},\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}},\ldots,\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}}}^{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\overbrace{0,1,\ldots,\RNW - 1}^{\text{$\RNW$ values}}}
        \end{align*}} \only<4>{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $0$}},\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}},\ldots,\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}}}^{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\overbrace{0,1,\ldots,\RNW - 1}^{\text{$\RNW$ values}}}
        \end{align*}} \only<5>{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $0$}},\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $n$}},\ldots,\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}}}^{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\overbrace{0,1,\ldots,\RNW - 1}^{\text{$\RNW$ values}}}
        \end{align*}} \only<6>{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $0$}},\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $n$}},\ldots,\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\substack{\text{mapped to} \\ 2^W - n - \RNW}}}^{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\overbrace{0,1,\ldots,\RNW - 1}^{\text{$\RNW$ values}}}
        \end{align*}} \only<7->{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $0$}},\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $n$}},\ldots,\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\substack{\text{mapped to} \\ 2^W - n - \RNW}}}^{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\underbrace{\overbrace{\BB{0},1,\ldots,\RNW - 1}^{\text{$\RNW$ values}}}_{\substack{\text{mapped to} \\ 2^W - \RNW}}}
        \end{align*}}
        \item<8-> Only numbers in the \ER{leftover} interval get mapped to $\ER{2^W - \RNW} > 2^W - n$
        \item<9-> Algorithm: \begin{itemize}
            \item[(1)]<10-> Draw $x \in [0,2^W)$ and compute $r = x \Mod n$
            \item[(2)]<11-> Return $r$ if $x - r > 2^W - n$ else goto \textbf{\GB{(1)}}
        \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}{The Java Algorithm}
    Algorithm: \begin{itemize}
        \item[(1)] Draw $x \in [0,2^W)$ and compute $r = x \Mod n$
        \item[(2)] Return $r$ if $x - r > 2^W - n$ else goto \textbf{\GB{(1)}}
    \end{itemize}
    \pause 
    \begin{block}{Efficiency}<+->
        \begin{itemize}[<+->]
            \item At least one integer division operation
            \item Number of integer divisions operations equal to number of \GB{rounds}
            \item Return a number in \GB{round} if $x < 2^W - \RNW$
            \item Happens with probability $\frac{2^W - \RNW}{2^W} > \frac{1}{2}$
            \item Expected number of integer division operations is $\frac{2^W}{2^W - \RNW} < 2$
        \end{itemize}
    \end{block}

\end{frame}

\subsection{The Fast-Dice-Roller Algorithm}
\begin{frame}{The Fast-Dice-Roller Algorithm}
    \pause

\end{frame}

\subsection{The Bitmask Algorithm}
\begin{frame}{The Bitmask Algorithm}
    \pause

\end{frame}