### 2024-02-25 Presentation/preliminaries.tex

\section{Preliminaries}

\subsection{Formal Definition}
\begin{frame}{Formal Definition}
    \pause
    \onslide<+->{\textbf{Input:}} \begin{itemize}[<+->]
        \item source of uniform random integers in $[0,2^W)$: \texttt{rand()}
        \item upper bound of interval $n \in \mathbb{N}$
    \end{itemize}
    \onslide<+->{\textbf{Output:} \begin{itemize}
        \item uniform random integer in interval $[0,n)$
    \end{itemize}}

    \vspace*{0.5cm}

    \begin{center}
        \begin{tikzpicture}[scale = 1.5]
            \draw<6->[time] (0,0) -- (9,0);
            \node<6->[time] at (0.015,-0.25) {$0$};

            \draw<7->[int, emorot] (5,0) -- (8,0);
            \node<7-> at (5.015,-0.25) {$\ER{a}$};
            \node<7-> at (8.0,-0.25) {$\ER{b}$};
            \node<7-> at (6.5,0.35) {$\ER{[a,b)}$};

            \draw<8->[int, goetheblau] (0,0) -- (3,0);
            \node<8-> at (3.0,-0.25) {$b - a$};
            \node<8-> at (1.5,0.35) {$[0,b - a)$};

            \draw<9->[time,black] (3.3,0.5) -- (4.7,0.5);
            \node<9-> at (4.0,0.72) {$\BB{+\,a}$};

            \node[opacity=0] at (4.5,1.0) {};
        \end{tikzpicture}
    \end{center}
\end{frame}

\subsection{Operations}
\begin{frame}{Operations}
    \pause 
    \begin{definition}[Common Operations]<+->
        \begin{itemize}[<+->]
            \item \makebox[4cm][l]{Integer-Division:} \makebox[1.5cm][l]{$x \div y$} $\coloneqq \lfloor\nicefrac{x}{y}\rfloor$
            \item \makebox[4cm][l]{Remainder-Operation:} \makebox[1.5cm][l]{$x \Mod y$} $\coloneqq x - (x \div y)y$
            \item \makebox[4cm][l]{Bit-\textsc{RightShift}:} \makebox[1.5cm][l]{$x \gg W$} $\coloneqq x \div 2^W$
            \item \makebox[4cm][l]{Bit-\textsc{LeftShift}:} \makebox[1.5cm][l]{$x \ll W$} $\coloneqq x \cdot 2^W$
            \item \makebox[4cm][l]{Bitwise-\textsc{And}:} \makebox[1.5cm][l]{$x \And y$} \onslide<+->{$\rightarrow x \Mod 2^W \coloneqq x \And (2^W - 1)$} 
        \end{itemize}
    \end{definition}

    \vspace*{0.5cm}

    \begin{definition}[Power Remainder]<+->
        For $W,n \in \mathbb{N}$, we write $\RNW$ for $2^W \Mod n$.
    \end{definition}

\end{frame}

\subsection{The Naive Approach}\label{sec:1.3}
\begin{frame}{The Naive Approach}
    \pause 
    \onslide<+->{\begin{center}
            \large{$\texttt{rand()} \Mod n$}
    \end{center}}

    \onslide<+->{Does this work?}
    \begin{itemize}[<+->]
        \item \GB{Yes}, the generated number is in $[0,n)$.
    \end{itemize}

    \onslide<+->{Is this efficient?}
    \begin{itemize}[<+->]
        \item \ER{No}, we require one expensive integer division operation.
    \end{itemize}

    \onslide<+->{Is the generated number uniform in $[0,n)$?}
\end{frame}

\begin{frame}{The Naive Approach - Bias}
    \pause 
    \onslide<+->{In general, applying $x \Mod n$ to $[0,2^W)$ yields}
    \onslide<+->{\begin{align*}\everymath{} % Allows specific coloring
        \overbrace{\underbrace{\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}},\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}},\ldots,\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\underbrace{0,1,\ldots,\RNW - 1}_{\text{$\RNW$ values}}}}^{\text{$2^W$ values}}
    \end{align*}}

    \onslide<+->{We have a \ER{leftover} interval that introduces bias.}

    \vspace*{0.5cm}

    \begin{block}{Deterministic Mappings}<+->
        \onslide<+->{\GB{Every} deterministic mapping $f\colon[0,2^W) \rightarrow [0,n)$}
        \onslide<+->{does \ER{not} generate \GB{uniform} random integers in one step}
        \onslide<+->{whenever $n$ does not divide $2^W$.}
    \end{block}
\end{frame}